Road Segment Prioritization for Bicycle Infrastructure

### 2020-12-11 README.Rmd
Mor updates

---
title: Road Segment Prioritization for Bicycle Infrastructure
date: "`r Sys.Date()`"
geometry: margin=2cm
bibliography:
  - mybibfile.bib
  - rlbibfile.bib
#linenumbers: true
numbersections: true
abstract: "Understanding the motivators and deterrents to cycling is essential for creating infrastructure that is successful in getting more people to adopt cycling as a mode of transport. The aim of this paper is to develop methods to prioritise cycling infrastructure, accounting for cyclist preference and UK government policies to create Low Traffic Neighborhoods. Distance decay, routing, and network analysis techniques are used to examine where future cycling demand is most likely to arise, how such demand could be accomodated by the street network, and how to ensure a fair distribution of investment. The city of Manchester, which has substantial cycling ambitions, is used to demonstrate the methods which, we demonstrate, can be applied to any city. Flexible routing profiles that are adjustable to local needs can avoid high-stress road segments, enabling the identification of networks that balance directness against the need for safe and stress-free routes. Community detection is used to partition the city and evaluate results from 'utilitarian' and 'egalitarian' policy perspectives. We conclude that the 'egalitarian' model prioritisation approach demonstrated in this paper has great potential to support cost-effective strategic cycle network planning internationally. \\par\\textbf{Keywords:} cycling networks, routing, transport equity"
# keywords: "cycling, low-traffic neighborhoods, potential demand, routing"
output:
  # bookdown::github_document2:
  bookdown::pdf_document2:
      toc: false
      extra_dependencies: ["subcaption", "longtable", "float", "multirow", "placeins", "multicol"]
---

@@

As noted by @buehler2016bikeway, it is only recently that research has pivoted from focusing on localized, street or intersection level impact towards studying the cycling network as a whole. They emphasize the promise of this shift in capturing the network-wide effect of street-level interventions. *Optimization* techniques have been used to propose improvements to cycling networks. @mesbah2012bilevel propose a bi-level formulation to optimize allocation of cycling lanes to the network without exceeding a set budget. The upper level is the proposed interventions and the lower level is the route choices made by users in reaction to changes in the network. The problem accounts for the effect of cycling lanes on car traffic, and attempts to maximize utilization of said lanes with minimal impact on car travel times. To improve cohesion of the suggested network, a constraint is added so that each link^[*link* refers to a road segment throughout this research] with a bike lane should be connected to at least one destination. Car usage is not considered by @mauttone2017bicycle, who develop an optimization framework that aims to minimize the total user cost of cycling on the network. The aggregate flow on links is obtained by using shortest paths to route existing cycling demand onto the road network, and the solution is a proposed set of links where cycling infrastructure should be added in order to minimize the overall travel cost of cyclists across the network. The cost of traversing a link is given as a function of its length and whether or not it has cycling infrastructure, and a discontinuity penalty is also added to prioritize connected road segments. The problem has also been solved by attempting to find the minimum cost of improving roadway links to meet a desired level of service (LOS) [@duthie2014optimization]. In this formulation, all OD pairs need to be connected by roads that meet the desired LOS, and a directness constraint is added so that paths between OD pairs do not exceed a certain multiple of the shortest path.

@@

The research provides a methodology for designing a cycling network that respects both the needs of the users and the ambitions of the government. The methodology can be used as a starting point for introducing a direct, connected, and low-stress network for any city by leveraging available open data. The only data necessary for reproducing it is the road network (from OSM), its topography (from satellite imagery), and commuter data (from the national census). It should be noted that the recommendations are only as good as the granularity of the available data.  
A core part of the methodology is determining where there is high potential for cycling, and using this as a basis for recommendations on road space reallocation in order to unlock potential cycling demand. A routing engine was used to study the effects of limiting access on different road types, and it was found that reducing cyclist flow on roads with high through-traffic resulted in acceptable increases in commuting distances. As a result, a hierarchy of road preference was used to route potential cycling demand. Algorithms were developed to determine where investments in cycling infrastructure should be prioritized. These algorithms, based on connectivity and egalitarian principles of resource distribution, ensured that whatever the level of investment, the resulting cycling network would improve connectivity of existing cycling infrastructure.

The research hopes to aid in the current shift towards active transport that is being promoted by the UK government. The success of such a shift is contingent on the appeal of the cycling network for prospective cyclists, a fact that was used as the basis for the methodology outlined here.

>>>>

---
title: Road Segment Prioritization for Bicycle Infrastructure
date: "`r Sys.Date()`"
geometry: margin=2cm
bibliography:
  - mybibfile.bib
  - rlbibfile.bib
#linenumbers: true
numbersections: true
abstract: "Understanding the motivators and deterrents to cycling is essential for creating infrastructure that is successful in getting more people to adopt cycling as a mode of transport. The aim of this paper is to develop methods to prioritise cycling infrastructure, accounting for cyclist preference and UK government policies to create Low Traffic Neighborhoods. Distance decay, routing, and network analysis techniques are used to examine where future cycling demand is most likely to arise, how such demand could be accomodated by the street network, and how to ensure a fair distribution of investment. The city of Manchester, which has substantial cycling ambitions, is used to demonstrate the methods which, we demonstrate, can be applied to any city. Flexible routing profiles that are adjustable to local needs can avoid high-stress road segments, enabling the identification of networks that balance directness against the need for safe and stress-free routes. Community detection is used to partition the city and evaluate results from 'utilitarian' and 'egalitarian' policy perspectives. We conclude that the 'egalitarian' model prioritisation approach demonstrated in this paper has great potential to support cost-effective strategic cycle network planning internationally. \\par\\textbf{Keywords:} cycling networks, routing, transport equity"
# keywords: "cycling, low-traffic neighborhoods, potential demand, routing"
output:
  bookdown::github_document2:
  # bookdown::pdf_document2:
      toc: false
      extra_dependencies: ["subcaption", "longtable", "float", "multirow", "placeins", "multicol"]
---

@@

As noted by @buehler2016bikeway, there has been shift in focus from localized (street or intersection) level planning towards planning based on studying the cycling network as a whole since the turn of the 21^st^ century. They emphasize the promise of this shift in capturing the network-wide effect of street-level interventions. *Optimization* techniques have been used to propose improvements to cycling networks. @mesbah2012bilevel propose a bi-level formulation to optimize allocation of cycling lanes to the network without exceeding a set budget. The upper level is the proposed interventions and the lower level is the route choices made by users in reaction to changes in the network. The problem accounts for the effect of cycling lanes on car traffic, and attempts to maximize utilization of said lanes with minimal impact on car travel times. To improve cohesion of the suggested network, a constraint is added so that each link^[*link* refers to a road segment throughout this research] with a bike lane should be connected to at least one destination. Car usage is not considered by @mauttone2017bicycle, who develop an optimization framework that aims to minimize the total user cost of cycling on the network. The aggregate flow on links is obtained by using shortest paths to route existing cycling demand onto the road network, and the solution is a proposed set of links where cycling infrastructure should be added in order to minimize the overall travel cost of cyclists across the network. The cost of traversing a link is given as a function of its length and whether or not it has cycling infrastructure, and a discontinuity penalty is also added to prioritize connected road segments. The problem has also been solved by attempting to find the minimum cost of improving roadway links to meet a desired level of service (LOS) [@duthie2014optimization]. In this formulation, all OD pairs need to be connected by roads that meet the desired LOS, and a directness constraint is added so that paths between OD pairs do not exceed a certain multiple of the shortest path.

@@

<!-- I think this section should start with a short (one sentence) re-statement of what the paper has done. -->

The research presented in this paper provides a methodology for designing a cycling network that respects both the needs of the users and the ambitions of transport planners and local/regional level government stakeholders. The methodology can be used as a starting point for introducing a direct, connected, and low-stress network for any city by leveraging available open data. The only data necessary for reproducing it is the road network (from OSM), its topography (from satellite imagery), and commuter data (from the national census). It should be noted that the recommendations are only as good as the granularity of the available data.

A core part of the methodology is determining where there is high potential for cycling, and using this as a basis for recommendations on road space reallocation in order to unlock potential cycling demand. A routing engine was used to study the effects of limiting access on different road types, and it was found that reducing cyclist flow on roads with high through-traffic resulted in acceptable increases in commuting distances. As a result, a hierarchy of road preference was used to route potential cycling demand. Algorithms were developed to determine where investments in cycling infrastructure should be prioritized. These algorithms, based on connectivity and egalitarian principles of resource distribution, ensured that whatever the level of investment, the resulting cycling network would improve connectivity of existing cycling infrastructure.

The methods can support the ongoing shift in transport planning towards active transport that is being promoted by the UK government and other authorities at city, regional and national levels worldwide. The success of such a shift depends on the appeal of the cycling network for prospective cyclists, a fact that was used as the basis for the methodology outlined here.

<!-- I think it would be good to say how others could reproduce the methods in other cities (RL) -->

<!-- There should also be links to the literature. -->
